---
  title: "CNV Analysis CPTAC-HNSC"
subtitle: "Obtatining Copy Number Index"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
  toc: true
theme: hpstr
highlight: github
number_sections: true
pdf_document:
  toc: true
toc_depth: 4
editor_options: 
  chunk_output_type: inline
bibliography: citations.bibtex
nocite: '@*'
---
  
# Obtain the Data with TCGABiolinks
  
## CNV data
  
This data was downloaded on the 2nd of March 2022 from the LinkedOmics portal: http://linkedomics.org/data_download/CPTAC-HNSCC/.

```{r}
Segment_CPTAC <- read.csv("/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/HS_CPTAC_HNSCC_SCNA_log2_segment_level.cvs", sep = "\t")
```

## Patient Metadata

This data was obtained on the 2nd of March from the GDC portal using TCGABiolinks function GDCquery_clinic:

```{r eval=FALSE}
clinical <- GDCquery_clinic(project = "CPTAC-3", type = "clinical")
clinical_SSC <- clinical[clinical$primary_diagnosis == "Squamous cell carcinoma, NOS",]
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Lower lobe, lung")
clinical_SSC <- clinical_SSC[i,]
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Lung, NOS")
clinical_SSC <- clinical_SSC[i,]
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Middle lobe, lung")
clinical_SSC <- clinical_SSC[i,]
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Upper lobe, lung")
clinical_SSC <- clinical_SSC[i,]
## Here you insert the pathway and filename you desire :)
write.csv(as.data.frame(clinical_SSC), 
          file="/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_clinical_HNSSC.csv")
i <- which(clinical_SSC$gender == "male")
clinical_SSC_male <- clinical_SSC[i,]
## Here you insert the pathway and filename you desire :)
write.csv(as.data.frame(clinical_SSC_male), 
          file="/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_clinical_HNSSC_male.csv")
```
  
```{r}
Segment_CPTAC <- read.csv("/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_clinical_HNSSC_male.csv", sep = "\t")
```

## Biospecimen data: Percentage Tumor Nuclei

This data was not found for this particular project.

# Processing of the data:

## Obtaining CNI

```{r}
HNSC_SegmentFiles <-Segment_CPTAC
```

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles$size <- HNSC_SegmentFiles$end - HNSC_SegmentFiles$start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles, f = HNSC_SegmentFiles$sample )
```

```{r}
duplicated(patients)
```

### In male patients

Selecting the male patients:

```{r}
SegmentFiles_Male <- HNSC_SegmentFiles[HNSC_SegmentFiles$sample %in% clinical_SSC_male$submitter_id,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Male$Unlog_Segment <- 2^(1+SegmentFiles_Male$log2)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Male$Weighted_value <- SegmentFiles_Male$Unlog_Segment*SegmentFiles_Male$size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Male, f = SegmentFiles_Male$sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
a <- paste0("chr", a)
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$chromosome %in% i])/sum(patients_male[[h]]$size[patients_male[[h]]$chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
chrom_male <- as.data.frame(apply( chrom_male, 2, as.numeric))
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

Plotting the CNI for all the Male patients:

```{r}
den <- density((chrom_male$chrY))
plot(den, xlim = c(0,2))
```

```{r}
den <- density((chrom_male$chrY/2))
plot(den, xlim = c(0,2))
```

Saving the data:

```{r}
write.csv(chrom_male, "/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/Male_CopyNumberIndex.csv")
```

Calculating the maximum points of the graph:

```{r}
which.max(den$y)
den$x[288]
```

```{r}
which.max(den$y[den$x < 1.917638])
den$x[72]
```

```{r}
den <- density(chrom_male$chrX)
plot(den)
```

```{r}
LoY <- chrom_male[chrom_male$chrY < 1,]
```

```{r}
i <- which(clinical_SSC_male$submitter_id %in% rownames(LoY))
i
```

```{r}
LoY <- c(1:length(clinical_SSC_male$submitter_id))
LoY[i] <- c("Yes")
LoY[-i] <- c("No")
```

```{r}
clinical_SSC_male$LoY <- LoY
```

```{r}
table(clinical_SSC_male$LoY)
```

```{r}
write.csv(clinical_SSC_male, "/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_clinical_HNSSC_male_LoY.csv")
```

### Doing the same process for the female patients:

```{r}
Female_ID <- TCGA_Masterfile[TCGA_Masterfile$Gender == "Female",]
```

```{r}
SegmentFiles_Female <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample %in% Female_ID$Patient_ID,]
```

```{r}
SegmentFiles_Female$Unlog_Segment <- 2^(1+SegmentFiles_Female$Segment_Mean)
```

```{r}
SegmentFiles_Female$Weighted_value <- SegmentFiles_Female$Unlog_Segment*SegmentFiles_Female$Size
```

```{r}
patients_Female <- split(SegmentFiles_Female, f = SegmentFiles_Female$Sample )
```

```{r}
ID <- names(patients_Female)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_Female))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_Female[[h]]$Weighted_value[patients_Female[[h]]$Chromosome %in% i])/sum(patients_Female[[h]]$Size[patients_Female[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

```{r}
chrom_Female <- do.call(rbind.data.frame, x)
```

```{r}
colnames(chrom_Female) <- a
rownames(chrom_Female) <- ID
```

```{r}
write.table(chrom_Female, "/Users/cristinacondel/Desktop/Gender_COX2/Female_CopyNumberIndex.txt", sep = "\t", row.names = T, quote = T)
```

```{r}
den <- density(chrom_Female$Y)
plot(den)
```

```{r}
den <- density(chrom_Female$X)
plot(den)
```
