---
title: "CNV Analysis TCGA-HNSC"
subtitle: "Obtatining Copy Number Index"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
 prettydoc::html_pretty:
    toc: true
    theme: hpstr
    highlight: github
    number_sections: true
 pdf_document:
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
bibliography: citations.bibtex
nocite: '@*'
---

# Obtain the Data with TCGABiolinks

## CNV data

This data was downloaded on the 31st of January 2022 from the GDC portal using R tool TCGABiolinks.

Loading of package TCGAbiolink (download raw data)

```{r, results= FALSE, message=FALSE, warning=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("TCGAbiolinks")
```

```{r, results= FALSE, message=FALSE, warning=FALSE}
library (TCGAbiolinks)
```


Specification of data search, we store the specifications in query 
Possible specifications (look in the GDC portal): GDCquery(
  project,
  data.category,
  data.type,
  workflow.type,
  legacy = FALSE,
  access,
  platform,
  file.type,
  barcode,
  data.format,
  experimental.strategy,
  sample.type
)

In this case selecting head and neck cancer, transcriptome profiling data --> gene expression quatification, data type in counts.

```{r, results= FALSE, message=FALSE}
query <- GDCquery(project = 'TCGA-HNSC',
data.category = "Copy Number Variation",
data.type = "Copy Number Segment",
platform = "Affymetrix SNP 6.0")
```

Obtaining sample names from the query file (column cases)

```{r}
samplesDown <- getResults (query, cols = c('cases'))
```

Selecting the samples that come from "Primary Solid Tumor" (typesample = "TP") among all the samples selected.

```{r}
dataSmTP <- TCGAquery_SampleTypes (barcode = samplesDown, typesample = 'NT')
```


Searching in the database for the selected NT samples:

```{r, message=FALSE, results=FALSE}
queryDown <- GDCquery(project = 'TCGA-HNSC',
data.category = "Copy Number Variation",
data.type = "Copy Number Segment",
platform = "Affymetrix SNP 6.0",
barcode = c(dataSmTP))
```

Once the search is defined then we obtain the data with GSCdownload, it downloads itself into a new directory in the working directory called GDC data, it can be changed in the function:

```{r, message=FALSE}
GDCdownload(queryDown)
```

GDCprepare converts the data downloaded into an R file

```{r, message=FALSE}
HNSC_SegmentFiles <- GDCprepare(queryDown, save = TRUE, save.filename =
"TCGA_HNSC_SegmentFiles.rda")
```

## Patient Metadata

This data was obtained on the 7th of October downloaded directly from the GDC portal: 

```{r}
clinical_data <- as.data.frame(SummarizedExperiment::colData(HNSC_SegmentFiles))
```

```{r}
write.csv(clinical_data, file = 'HNSC_NT_Clinical_data.csv')
```

```{r}
male_ID <- clinical_data[ clinical_data$gender == "male",]
```

# Processing of the data:

## Obtaining CNI

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles$Size <- HNSC_SegmentFiles$End - HNSC_SegmentFiles$Start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles, f = HNSC_SegmentFiles$Sample )
```

Giving patients the same ID as other datasets:

```{r}
name <- names(patients)
```

```{r}
name <- substring(name, 1, 12)
```


```{r}
duplicated(name) 
```

Taking out duplicated data:

Renaming the patients with an ID matching the metadata:

```{r}
HNSC_SegmentFiles$Sample <- substring(HNSC_SegmentFiles$Sample, 1, 12)
```

### In male patients

Selecting the male patients:

```{r}
i <- which(HNSC_SegmentFiles$Sample %in% male_ID$patient)
```


```{r}
SegmentFiles_Male <- HNSC_SegmentFiles[i,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Male$Unlog_Segment <- 2^(1+SegmentFiles_Male$Segment_Mean)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Male$Weighted_value <- SegmentFiles_Male$Unlog_Segment*SegmentFiles_Male$Size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Male, f = SegmentFiles_Male$Sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$Chromosome %in% i])/sum(patients_male[[h]]$Size[patients_male[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

Plotting the CNI for all the Male patients:

```{r}
den <- density(chrom_male$Y)
png(file= "plotY_TCGA_HNSC_NT.png", width=600, height=350)
plot(den, xlim = c(0,2))
dev.off()
```




