---
title: "CNV Analysis TCGA-HNSC"
subtitle: "Obtatining Copy Number Index"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
 prettydoc::html_pretty:
    toc: true
    theme: hpstr
    highlight: github
    number_sections: true
 pdf_document:
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
bibliography: citations.bibtex
nocite: '@*'
---

# Obtain the Data with TCGABiolinks

## CNV data

This data was downloaded on the 31st of January 2022 from the GDC portal using R tool TCGABiolinks.

Loading of package TCGAbiolink (download raw data)

```{r, results= FALSE, message=FALSE, warning=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("TCGAbiolinks")
```

```{r, results= FALSE, message=FALSE, warning=FALSE}
library (TCGAbiolinks)
```


Specification of data search, we store the specifications in query 
Possible specifications (look in the GDC portal): GDCquery(
  project,
  data.category,
  data.type,
  workflow.type,
  legacy = FALSE,
  access,
  platform,
  file.type,
  barcode,
  data.format,
  experimental.strategy,
  sample.type
)

In this case selecting head and neck cancer, transcriptome profiling data --> gene expression quatification, data type in counts.

```{r, results= FALSE, message=FALSE}
query <- GDCquery(project = 'TCGA-HNSC',
data.category = "Copy Number Variation",
data.type = "Copy Number Segment",
platform = "Affymetrix SNP 6.0")
```

Obtaining sample names from the query file (column cases)

```{r}
samplesDown <- getResults (query, cols = c('cases'))
```

Selecting the samples that come from "Primary Solid Tumor" (typesample = "TP") among all the samples selected.

```{r}
dataSmTP <- TCGAquery_SampleTypes (barcode = samplesDown, typesample = 'TP')
```


Searching in the database for the selected TP samples:

```{r, message=FALSE, results=FALSE}
queryDown <- GDCquery(project = 'TCGA-HNSC',
data.category = "Copy Number Variation",
data.type = "Copy Number Segment",
platform = "Affymetrix SNP 6.0",
barcode = c(dataSmTP))
```

Once the search is defined then we obtain the data with GSCdownload, it downloads itself into a new directory in the working directory called GDC data, it can be changed in the function:

```{r, message=FALSE}
GDCdownload(queryDown)
```

GDCprepare converts the data downloaded into an R file

```{r, message=FALSE}
HNSC_SegmentFiles <- GDCprepare(queryDown, save = TRUE, save.filename =
"TCGA_HNSC_SegmentFiles.rda")
```

## Patient Metadata

This data was obtained on the 7th of October downloaded directly from the GDC portal: 

```{r, warning=FALSE}
library(readxl)
TCGA_Masterfile <- read_excel("TCGA_MASTERFILE_clinical_data.xls")
TCGA_Masterfile$Patient_ID <- gsub("-", ".", TCGA_Masterfile$Patient_ID)
row.names(TCGA_Masterfile) <- TCGA_Masterfile$Patient_ID
```

```{r}
male_ID <- TCGA_Masterfile[TCGA_Masterfile$Gender=="Male",]
```

## Biospecimen data: Percentage Tumor Nuclei

This data was downloaded on the 2nd of February 2022 directly from the GDC portal, the same parameters for selecting relevant patients were used as in the data download above.

We read the data:

```{r}
orgs <- read.delim("slide.tsv", header=TRUE)
```

We take the relevant information from the dataset, in this case Patient ID and percentage of tumor cell nuclei in the sample.

```{r}
data <- data.frame(ID = orgs$case_submitter_id, perc_nuc = orgs$percent_tumor_nuclei)
```

We transform the ID's into the same format as other datasets:

```{r}
data$ID <- gsub("-", ".", data$ID)
```

```{r}
data <- data[data$ID  %in% male_ID$Patient_ID,]
```

```{r, warning=FALSE}
data1 <- data[!is.na(as.numeric(data$perc_nuc)), ]
```

```{r}
data2 <- aggregate(as.numeric(data1$perc_nuc),by=list(ID=data1$ID),data=data1,FUN=mean)
```

# Processing of the data:

## Obtaining CNI

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles$Size <- HNSC_SegmentFiles$End - HNSC_SegmentFiles$Start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles, f = HNSC_SegmentFiles$Sample )
```

Giving patients the same ID as other datasets:

```{r}
name <- names(patients)
```

```{r}
name <- substring(name, 1, 12)
```

```{r}
name <- gsub("-", ".", name)
```

```{r}
duplicated(name) 
```

Taking out duplicated data:

```{r}
HNSC_SegmentFiles <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample != "TCGA-CV-6955-01A-11D-2062-01",]
```

```{r}
HNSC_SegmentFiles <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample != "TCGA-CV-6956-01A-21D-2062-01",]
```

Renaming the patients with an ID matching the metadata:

```{r}
HNSC_SegmentFiles$Sample <- substring(HNSC_SegmentFiles$Sample, 1, 12)
HNSC_SegmentFiles$Sample <- gsub("-", ".", HNSC_SegmentFiles$Sample)
```

### In male patients

Selecting the male patients:

```{r}
SegmentFiles_Male <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample %in% male_ID$Patient_ID,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Male$Unlog_Segment <- 2^(1+SegmentFiles_Male$Segment_Mean)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Male$Weighted_value <- SegmentFiles_Male$Unlog_Segment*SegmentFiles_Male$Size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Male, f = SegmentFiles_Male$Sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$Chromosome %in% i])/sum(patients_male[[h]]$Size[patients_male[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

Plotting the CNI for all the Male patients:

```{r}
den <- density(chrom_male$Y)
plot(den, xlim = c(0,2))
```
Obtaining the data in common patients for Percentage Tumor Nuclei and Chromosome Number Index:

```{r}
data3 <- data2[data2$ID %in% rownames(chrom_male),]
```

```{r}
chrom_male1 <- chrom_male[data3$ID,]
```

Weighting Chromosome Number Index by Percentage tumor nuclei:

```{r}
chrom_male2 <- (chrom_male1 + (data3$x/100) - 1)/(data3$x/100)
```

Saving the data:

```{r}
write.table(chrom_male2, "Male_CopyNumberIndex.txt", sep = "\t", row.names = F, quote = F)
```

Plotting weighted Y chromosome for male patients:

```{r}
den <- density(chrom_male2$`Y`)
plot(den, xlim = c(0,2), xlab = "ChrY Copy Number Index")
```

From this graph we can extract that we have successfully replicated what was done in the analysis by Hollows et al. 

Calculating the maximum points of the graph:

```{r}
which.max(den$y)
den$x[110]
```

```{r}
which.max(den$y[den$x < 0.75])
den$x[72]
```

```{r}
den <- density(chrom_male2$`X`)
plot(den)
```

```{r}
LoY <- chrom_male2[chrom_male2$Y < 0.4803115,]
```

```{r}
i <- which(male_ID$Patient_ID %in% rownames(LoY))
i
```

```{r}
LoY <- c(1:length(male_ID$Patient_ID))
LoY[i] <- c("Yes")
LoY[-i] <- c("No")
```

```{r}
male_ID$LoY <- LoY
```


### Doing the same process for the female patients:

```{r}
Female_ID <- TCGA_Masterfile[TCGA_Masterfile$Gender == "Female",]
```

```{r}
SegmentFiles_Female <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample %in% Female_ID$Patient_ID,]
```

```{r}
SegmentFiles_Female$Unlog_Segment <- 2^(1+SegmentFiles_Female$Segment_Mean)
```

```{r}
SegmentFiles_Female$Weighted_value <- SegmentFiles_Female$Unlog_Segment*SegmentFiles_Female$Size
```

```{r}
patients_Female <- split(SegmentFiles_Female, f = SegmentFiles_Female$Sample )
```

```{r}
ID <- names(patients_Female)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_Female))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_Female[[h]]$Weighted_value[patients_Female[[h]]$Chromosome %in% i])/sum(patients_Female[[h]]$Size[patients_Female[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

```{r}
chrom_Female <- do.call(rbind.data.frame, x)
```

```{r}
colnames(chrom_Female) <- a
rownames(chrom_Female) <- ID
```

```{r}
write.table(chrom_Female, "Female_CopyNumberIndex.txt", sep = "\t", row.names = T, quote = T)
```

```{r}
den <- density(chrom_Female$Y)
plot(den, xlim = c(0,2), xlab = "ChrY Copy Number Index")
```

```{r}
den <- density(chrom_Female$X)
plot(den)
```

# Bibliography

