---
title: "Untitled"
output: html_document
date: '2022-08-17'
---

```{r}
Segment_CPTAC <- read.csv("HS_CPTAC_HNSCC_SCNA_log2_segment_level.cvs", sep = "\t")
HNSC_SegmentFiles <- Segment_CPTAC
```

```{r}
clinical_SSC_male <- read.csv("CPTAC_clinical_HNSSC_male.csv", sep = "\t")
```


Extract all the measures for the Y chromosome:

```{r}
HNSC_SegmentFiles_Y <- HNSC_SegmentFiles[HNSC_SegmentFiles$chromosome == "chrY",]
```

Divide the measures taking into consideration if they belong to the p or q arm:

```{r}
HNSC_SegmentFiles_Y_p <- HNSC_SegmentFiles_Y[HNSC_SegmentFiles_Y$start > 1 & HNSC_SegmentFiles_Y$end < 10400000, ]
```

```{r}
HNSC_SegmentFiles_Y_q <- HNSC_SegmentFiles_Y[HNSC_SegmentFiles_Y$start > 10400001 & HNSC_SegmentFiles_Y$end < 57227415, ]
```

Calculating chromosome index for these regions:

# Processing of the data:

## Obtaining CNI

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles_Y_p$Size <- HNSC_SegmentFiles_Y_p$end - HNSC_SegmentFiles_Y_p$start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles_Y_p, f = HNSC_SegmentFiles_Y_p$sample )
```

Giving patients the same ID as other datasets:

```{r}
name <- names(patients)
```

```{r}
duplicated(name) 
```

### In male patients

Selecting the male patients:

```{r}
SegmentFiles_Y_p <- HNSC_SegmentFiles_Y_p[HNSC_SegmentFiles_Y_p$sample %in% CPTAC_clinical_HNSSC_male$submitter_id...2,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Y_p$Unlog_Segment <- 2^(1+SegmentFiles_Y_p$log2)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Y_p$Weighted_value <- SegmentFiles_Y_p$Unlog_Segment*SegmentFiles_Y_p$Size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Y_p, f = SegmentFiles_Y_p$sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
a <- paste0("chr", a)
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$chromosome %in% i])/sum(patients_male[[h]]$Size[patients_male[[h]]$chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

```{r}
chrom_male <- chrom_male[,c(-(1:23))]
```

```{r}
chrom_male <- as.data.frame(chrom_male)
rownames(chrom_male) <- ID
colnames(chrom_male) <- c("Y_p")
```



Plotting the CNI for all the Male patients:

```{r}
den <- density(chrom_male$Y_p)
plot(den)
```

```{r}
chrom_male$ID <- rownames(chrom_male) 
```

```{r}
i <- which(chrom_male$Y_p < 1)
P_Loss <- as.data.frame(chrom_male[i,])
rownames(P_Loss) <- rownames(chrom_male)[i]
```

```{r}
i <- which(CPTAC_clinical_HNSSC_male$submitter_id...2 %in% rownames(P_Loss))
i
```

```{r}
P_Loss <- c(1:length(CPTAC_clinical_HNSSC_male$submitter_id...2))
P_Loss[i] <- c("Yes")
P_Loss[-i] <- c("No")
```

```{r}
P_Loss <- as.data.frame(P_Loss)
rownames(P_Loss) <- CPTAC_clinical_HNSSC_male$submitter_id...2
```

```{r}
male_ID$P_Loss <- P_Loss
```

# Processing of the data:

## Obtaining CNI

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles_Y_q$Size <- HNSC_SegmentFiles_Y_q$end - HNSC_SegmentFiles_Y_q$start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles_Y_q, f = HNSC_SegmentFiles_Y_q$sample )
```

Giving patients the same ID as other datasets:

```{r}
name <- names(patients)
```

```{r}
duplicated(name) 
```

### In male patients

Selecting the male patients:

```{r}
SegmentFiles_Y_q <- HNSC_SegmentFiles_Y_q[HNSC_SegmentFiles_Y_q$sample %in% CPTAC_clinical_HNSSC_male$submitter_id...2,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Y_q$Unlog_Segment <- 2^(1+SegmentFiles_Y_q$log2)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Y_q$Weighted_value <- SegmentFiles_Y_q$Unlog_Segment*SegmentFiles_Y_q$Size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Y_q, f = SegmentFiles_Y_q$sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
a <- paste0("chr", a)
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$chromosome %in% i])/sum(patients_male[[h]]$Size[patients_male[[h]]$chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

```{r}
chrom_male <- chrom_male[,c(-(1:23))]
```

```{r}
chrom_male <- as.data.frame(chrom_male)
rownames(chrom_male) <- ID
colnames(chrom_male) <- c("Y_q")
```



Plotting the CNI for all the Male patients:

```{r}
den <- density(chrom_male$Y_q)
plot(den)
```


```{r}
i <- which(chrom_male$Y_q < 1)
Q_Loss <- as.data.frame(chrom_male[i,])
rownames(Q_Loss) <- rownames(chrom_male)[i]
```

```{r}
i <- which(CPTAC_clinical_HNSSC_male$submitter_id...2 %in% rownames(Q_Loss))
i
```

```{r}
Q_Loss <- c(1:length(CPTAC_clinical_HNSSC_male$submitter_id...2))
Q_Loss[i] <- c("Yes")
Q_Loss[-i] <- c("No")
```

```{r}
Partial_loss <- P_Loss
Partial_loss$Y_q <- Q_Loss
colnames(Partial_loss)[1] <- "Y_p"
```

```{r}
Partial_loss$LoY <- CPTAC_clinical_HNSSC_male_LoY$LoY
```


```{r}
write_csv(Partial_loss, "Q_P_Loss.csv")
```