---
title: "Extracting data with TCGABiolinks"
output: html_document
---

Loading of package TCGAbiolink (download raw data)

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("TCGAbiolinks")
```




Specification of data search, we store the specifications in query 
Possible specifications (look in the GDC portal): GDCquery(
  project,
  data.category,
  data.type,
  workflow.type,
  legacy = FALSE,
  access,
  platform,
  file.type,
  barcode,
  data.format,
  experimental.strategy,
  sample.type
)

In this case selecting head and neck cancer, transcriptome profiling data --> gene expression quatification, data type in counts.

```{r}
query <- GDCquery(project = 'CPTAC-3',
data.category = "Transcriptome Profiling",
data.type = "Gene Expression Quantification",
workflow.type = "HTSeq - Counts", sample.type = "Primary Tumor")
```

Obtaining sample names from the query file (column cases)

```{r}
samplesDown <- getResults (query, cols = c('cases'))
```

Selecting the samples that come from "Primary Solid Tumor" (typesample = "TP") among all the samples selected.

```{r}
dataSmTP <- TCGAquery_SampleTypes (barcode = samplesDown, typesample = 'Primary Tumor')
```


Searching in the database for the selected TP samples:

```{r}
queryDown <- GDCquery(project = 'CPTAC-3',
data.category = "Transcriptome Profiling",
data.type = "Gene Expression Quantification",
workflow.type = "HTSeq - Counts",
barcode = c(dataSmTP), sa)
```

Once the search is defined then we obtain the data with GSCdownload, it downloads itself into a new directory in the working directory called GDC data, it can be changed in the function:

```{r}
GDCdownload(query)
```

GDCprepare converts the data downloaded into an R file

```{r}
CPTAC_HTseqSE <- GDCprepare(query, save = TRUE, save.filename =
"CPTAC_3_HTseq.rda")
```

Extracting the counts (under assay) using SummarizedExperiment from the HNSC_HTseqSE object

```{r}
Counts <- as.data.frame(SummarizedExperiment::assay(CPTAC_HTseqSE))
```

Writing the table into the working directory:

```{r}
write.table (Counts, file = 'HTseq2.tsv')
```

Renaming the columns so they have matching names

```{r}
colnames <- substr(colnames(Counts), 1, 12)
colnames(Counts) <- colnames
head(Counts)
```

Ordering the count table in the same alphanumeric way as the sample information table

```{r}
Counts <- Counts[ , order(names(Counts))]
```

Writing table in csv format into the working directory so its easier to get it in further uses

```{r}
## Here you insert the pathway and filename you desire :)

write.csv(as.data.frame(Counts), 
          file="/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_3_counts.csv")
```

```{r}
clinical <- GDCquery_clinic(project = "CPTAC-3", type = "clinical")
```

```{r}
clinical_SSC <- clinical[clinical$primary_diagnosis == "Squamous cell carcinoma, NOS",]
```

```{r}
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Lower lobe, lung")
clinical_SSC <- clinical_SSC[i,]
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Lung, NOS")
clinical_SSC <- clinical_SSC[i,]
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Middle lobe, lung")
clinical_SSC <- clinical_SSC[i,]
i <- which(clinical_SSC$tissue_or_organ_of_origin != "Upper lobe, lung")
clinical_SSC <- clinical_SSC[i,]
```

```{r}
## Here you insert the pathway and filename you desire :)

write.csv(as.data.frame(clinical_SSC), 
          file="/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC_clinical_HNSSC.csv")
```


```{r}
i <- which(clinical_SSC$gender == "male")
clinical_SSC_male <- clinical_SSC[i,]
```

```{r}
## Here you insert the pathway and filename you desire :)

write.csv(as.data.frame(clinical_SSC_male), 
          file="/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC_clinical_HNSSC_male.csv")
```

```{r}
names <- substr(colnames(Counts), 1, 9)
```

```{r}
colnames(Counts) <- names
```

```{r}
i <- which(colnames(Counts) %in% clinical_SSC_male$submitter_id)
```

```{r}
Counts_male <- Counts[,i]
```

```{r}
table((colnames(Counts) %in% clinical_SSC$submitter_id))
```






