---
title: "Untitled"
output: html_document
date: '2022-08-17'
---

Extract all the measures for the Y chromosome:

```{r}
HNSC_SegmentFiles_Y <- HNSC_SegmentFiles[HNSC_SegmentFiles$Chromosome == "Y",]
```

Divide the measures taking into consideration if they belong to the p or q arm:

```{r}
HNSC_SegmentFiles_Y_p <- HNSC_SegmentFiles_Y[HNSC_SegmentFiles_Y$Start > 1 & HNSC_SegmentFiles_Y$End < 10400000, ]
```

```{r}
HNSC_SegmentFiles_Y_q <- HNSC_SegmentFiles_Y[HNSC_SegmentFiles_Y$Start > 10400001 & HNSC_SegmentFiles_Y$End < 57227415, ]
```

Calculating chromosome index for these regions:

# Processing of the data:

## Obtaining CNI

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles_Y_p$Size <- HNSC_SegmentFiles_Y_p$End - HNSC_SegmentFiles_Y_p$Start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles_Y_p, f = HNSC_SegmentFiles_Y_p$Sample )
```

Giving patients the same ID as other datasets:

```{r}
name <- names(patients)
```

```{r}
name <- substring(name, 1, 12)
```

```{r}
name <- gsub("-", ".", name)
```

```{r}
duplicated(name) 
```

```{r}
name[164]
```


Taking out duplicated data:

```{r}
HNSC_SegmentFiles_Y_p <- HNSC_SegmentFiles_Y_p[HNSC_SegmentFiles_Y_p$Sample != "TCGA-CV-6955-01A-11D-2062-01",]
```

Renaming the patients with an ID matching the metadata:

```{r}
HNSC_SegmentFiles_Y_p$Sample <- substring(HNSC_SegmentFiles_Y_p$Sample, 1, 12)
HNSC_SegmentFiles_Y_p$Sample <- gsub("-", ".", HNSC_SegmentFiles_Y_p$Sample)
```

### In male patients

Selecting the male patients:

```{r}
SegmentFiles_Y_p <- HNSC_SegmentFiles_Y_p[HNSC_SegmentFiles_Y_p$Sample %in% male_ID$Patient_ID,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Y_p$Unlog_Segment <- 2^(1+SegmentFiles_Y_p$Segment_Mean)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Y_p$Weighted_value <- SegmentFiles_Y_p$Unlog_Segment*SegmentFiles_Y_p$Size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Y_p, f = SegmentFiles_Y_p$Sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$Chromosome %in% i])/sum(patients_male[[h]]$Size[patients_male[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

```{r}
chrom_male <- chrom_male[,c(-(1:23))]
```

```{r}
chrom_male <- as.data.frame(chrom_male)
rownames(chrom_male) <- ID
colnames(chrom_male) <- c("Y_p")
```



Plotting the CNI for all the Male patients:

```{r}
den <- density(chrom_male$Y_p)
plot(den)
```
```{r}
which.max(den$y)
den$x[108]
```


```{r}
which.max(den$y[den$x < 0.75])
den$x[72]
```

```{r}
chrom_male$ID <- rownames(chrom_male) 
```

```{r}
P_Loss <- as.data.frame(chrom_male[chrom_male$Y_p < 0.6321037,])
```

```{r}
i <- which(male_ID$Patient_ID %in% rownames(P_Loss))
i
```

```{r}
P_Loss <- c(1:length(male_ID$Patient_ID))
P_Loss[i] <- c("Yes")
P_Loss[-i] <- c("No")
```

```{r}
male_ID$P_Loss <- P_Loss
```

# Processing of the data:

## Obtaining CNI

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles_Y_q$Size <- HNSC_SegmentFiles_Y_q$End - HNSC_SegmentFiles_Y_q$Start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles_Y_q, f = HNSC_SegmentFiles_Y_q$Sample )
```

Giving patients the same ID as other datasets:

```{r}
name <- names(patients)
```

```{r}
name <- substring(name, 1, 12)
```

```{r}
name <- gsub("-", ".", name)
```

```{r}
duplicated(name) 
```

```{r}
name[173]
```


Taking out duplicated data:

```{r}
HNSC_SegmentFiles_Y_q <- HNSC_SegmentFiles_Y_q[HNSC_SegmentFiles_Y_q$Sample != "TCGA-CV-6955-01A-11D-2062-01",]
```

Renaming the patients with an ID matching the metadata:

```{r}
HNSC_SegmentFiles_Y_q$Sample <- substring(HNSC_SegmentFiles_Y_q$Sample, 1, 12)
HNSC_SegmentFiles_Y_q$Sample <- gsub("-", ".", HNSC_SegmentFiles_Y_q$Sample)
```

### In male patients

Selecting the male patients:

```{r}
SegmentFiles_Y_q <- HNSC_SegmentFiles_Y_q[HNSC_SegmentFiles_Y_q$Sample %in% male_ID$Patient_ID,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Y_q$Unlog_Segment <- 2^(1+SegmentFiles_Y_q$Segment_Mean)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Y_q$Weighted_value <- SegmentFiles_Y_q$Unlog_Segment*SegmentFiles_Y_q$Size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Y_q, f = SegmentFiles_Y_q$Sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$Chromosome %in% i])/sum(patients_male[[h]]$Size[patients_male[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

```{r}
chrom_male <- chrom_male[,c(-(1:23))]
```

```{r}
chrom_male <- as.data.frame(chrom_male)
rownames(chrom_male) <- ID
colnames(chrom_male) <- c("Y_q")
```



Plotting the CNI for all the Male patients:

```{r}
den <- density(chrom_male$Y_q)
plot(den)
```
```{r}
which.max(den$y)
den$x[108]
```


```{r}
which.max(den$y[den$x < 0.75])
den$x[111]
```

```{r}
chrom_male$ID <- rownames(chrom_male) 
```

```{r}
Q_Loss <- as.data.frame(chrom_male[chrom_male$Y_q < 0.572527,])
```

```{r}
i <- which(male_ID$Patient_ID %in% rownames(Q_Loss))
i
```

```{r}
Q_Loss <- c(1:length(male_ID$Patient_ID))
Q_Loss[i] <- c("Yes")
Q_Loss[-i] <- c("No")
```

```{r}
male_ID$Q_Loss <- Q_Loss
```


```{r}
table(male_ID$P_Loss, male_ID$Q_Loss)
```

```{r}
write_csv(male_ID, "Q_P_Loss.csv")
```
