---
title: "CNV Analysis TCGA Multiple Projects (Male > 50)"
subtitle: "Obtatining Copy Number Index"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
 prettydoc::html_pretty:
    toc: true
    theme: hpstr
    highlight: github
    number_sections: true
 pdf_document:
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
bibliography: citations.bibtex
nocite: '@*'
---

```{r, message=FALSE, eval=FALSE}
library(rlist)
```
This in case there is tumor nuclei percent information:

```{r, eval=FALSE}
require(data.table)
biospecimen <-as.data.frame(fread('slide.tsv'))
df <- lapply(biospecimen, function(x) as.data.frame(gsub("'--", NA, x)))
data_frame <- data.frame(df$case_submitter_id, df$percent_tumor_nuclei)
```


# Obtaining CNI

Upload 'TCGA_onlymalemore50_segmented_data.rdata', which contains the Segmented Data for the males only into h variable:

```{r, eval=FALSE}
h <- x
```

The Segmented Data is introduced in a loop prepared for the calculation of the Copy Number Index:

```{r, eval=FALSE}
f <- list()
for (proj in name_proj) {
  # Extracting the dataset with Segmented Data for the specific project:
  TCGA_BLCA <- do.call(rbind.data.frame, h[proj])
  # Calculating all the variables necessary for CNI:
  TCGA_BLCA$Size <- TCGA_BLCA$End - TCGA_BLCA$Start
  TCGA_BLCA$Unlog_Segment <- 2^(1+TCGA_BLCA$Segment_Mean)
  TCGA_BLCA$Weighted_value <- TCGA_BLCA$Unlog_Segment*TCGA_BLCA$Size
  patients_male <- split(TCGA_BLCA, f = TCGA_BLCA$Sample)
  ID <- names(patients_male)
  a <- c(1:22, "X", "Y")
  # Loop for calculating the CNI:
  x <- list()
    for (v in (1:length(patients_male))) {
      vec <- c()
      for (i in a) {
        mean <- sum(patients_male[[v]]$Weighted_value[patients_male[[v]]$Chromosome %in% i])/sum(patients_male[[v]]$Size[patients_male[[v]]$Chromosome %in% i])
        vec <- c(vec,mean)
      }
    names(vec) <- c(1:22, "X", "Y")
    x[[paste0(ID[v])]] <- vec
    }
  # Unfolding the list with the results in a dataframe, and naming of the colums and rows:
  chrom_male <- do.call(rbind.data.frame, x)
  colnames(chrom_male) <- a
  rownames(chrom_male) <- ID
  # Adding each new dataframe to a list, each object each project:
  f[[paste0(proj)]] <- chrom_male
}
```

A list of datasets with CNI Data for each project is generated and stored as a rdata object:

```{r, eval=FALSE}
list.save(f, 'TCGA_male_chrom_index.rdata')
```

# Plotting the Chromosome Index for the Y and X chromosomes in Male patients

The CNI Data is introduced in a loop prepared to plot the density of CNI of the Y chromosome and save the plots in the directory of this file:

```{r, eval=FALSE}
for (proj in name_proj) {
  TCGA_BLCA <- do.call(rbind.data.frame, f[proj])
  den <- density(TCGA_BLCA$Y)
  file <- paste0("plotY_", proj, ".png")
  png(file= file, width=600, height=350)
  title <- paste0("Chrom Y in Males ", proj)
  plot(den, xlim = c(0,2), main = title) 
  dev.off()
}
```

The CNI Data is introduced in a loop prepared to plot the density of CNI of the X chromosome and save the plots in the directory of this file:

```{r, eval=FALSE}
for (proj in name_proj) {
  TCGA_BLCA <- do.call(rbind.data.frame, x[proj])
  den <- density(TCGA_BLCA$X)
  file <- paste0("plotX_", proj, ".png")
  png(file= file, width=600, height=350)
  title <- paste0("Chrom X in Males ", proj)
  plot(den, xlim = c(0,4), main = title) 
  dev.off()
}
```

The CNI for the TCGA-PCPG cohort is not inside the parameters that all the others are (main peak around 4), we decide to take out this project:

```{r, eval=FALSE}
 TCGA_BLCA <- do.call(rbind.data.frame, x["TCGA-PCPG"])
  den <- density(TCGA_BLCA$Y)
  title <- paste0("Chrom Y in Males ","TCGA-PCPG")
  plot(den, main = title) 

```


```{r, eval=FALSE}
name_proj <- names(f)
#Taking out the "TCGA-PCPG" cohort since the values for the Y chromosome index are not okay
name_proj <- name_proj[-14]
```

The CNI Data is now introduced in a loop prepared to calculate the LoY status of each patient, in this case we have used 0.5 as the threshold under which a patient is considered to have a Loss of the Y chromosome:

```{r, eval=FALSE}
hs <- list()
for (proj in name_proj) {
TCGA_BLCA <- do.call(rbind.data.frame, f[proj])
  name <- rownames(TCGA_BLCA)
  name <- substr(name,11,22)
  rownames(TCGA_BLCA) <- name
  LoY1 <- TCGA_BLCA[TCGA_BLCA$Y < 0.5,]
  i <- which(rownames(TCGA_BLCA) %in% rownames(LoY1))
  LoY <- c(1:length(rownames(TCGA_BLCA)))
  LoY[i] <- c("Yes")
  LoY[-i] <- c("No")
  LoY <- as.data.frame(LoY)
  colnames(LoY) <- c("LoY")
  rownames(LoY) <- name
  hs[[paste0(proj)]] <- LoY
}
```

The only cohort that has no "Yes" value for LoY among all the cohorts its TCGA-THYM, the loop created is not working for cohorts in which this happens so we would replace this manually:

```{r}
hs <- hs[-21]
TCGA_BLCA <- do.call(rbind.data.frame, x[22])
  name <- rownames(TCGA_BLCA)
  name <- substr(name,11,22)
  rownames(TCGA_BLCA) <- name
  LoY <- replicate(length(rownames(TCGA_BLCA)), "No")
  LoY <- as.data.frame(LoY)
  colnames(LoY) <- c("LoY")
  rownames(LoY) <- name
```

```{r}
hs[["TCGA-THYM"]] <- LoY
```

A list of datasets with LoY Status Data for each project is generated and stored as a rdata object:

```{r}
list.save(hs, 'TCGA_onlymalemore50_LoYstatus.rdata')
```

```{r}
hs <- list()
for (proj in name_proj) {
TCGA_BLCA <- do.call(rbind.data.frame, hs[proj])
  LoY <- TCGA_BLCA[TCGA_BLCA$Y < 0.5,]
  name <- rownames(LoY)
  name <- substr(name,11,22)
  rownames(LoY) <- name
  clinical <- do.call(rbind.data.frame, projects[proj])
  i <- which(duplicated(clinical$case_submitter_id))
  clinical <- clinical[-i,]
  i <- which(clinical$case_submitter_id %in% rownames(LoY))
  LoY <- c(1:length(clinical$case_submitter_id))
  LoY[i] <- c("Yes")
  LoY[-i] <- c("No")
  clinical$LoY <- LoY
  hs[[paste0(proj)]] <- clinical
}

```


```{r}
hs <- list()
for (proj in name_proj) {
TCGA_BLCA <- do.call(rbind.data.frame, x[proj])
  LoY <- TCGA_BLCA[TCGA_BLCA$Y < 0.5,]
  name <- rownames(LoY)
  name <- substr(name,11,22)
  rownames(LoY) <- name
  clinical <- do.call(rbind.data.frame, projects[proj])
  i <- which(duplicated(clinical$case_submitter_id))
  clinical <- clinical[-i,]
  i <- which(clinical$case_submitter_id %in% rownames(LoY))
  LoY <- c(1:length(clinical$case_submitter_id))
  LoY[i] <- c("Yes")
  LoY[-i] <- c("No")
  clinical$LoY <- LoY
  hs[[paste0(proj)]] <- clinical
}

```

```{r}
list.save(hs, 'TCGA_onlymalemore50_clinicalndLoYstatus.rdata')
```

```{r}
clinicaldata <- do.call(rbind.data.frame, x)
```

```{r}
rownames(clinicaldata) <- clinicaldata$case_submitter_id
```

```{r}
clinicaldata <- clinicaldata[sort(rownames(clinicaldata)),]
```

```{r}
countdata <- do.call(cbind.data.frame, x)
```


```{r}
name <- colnames(countdata)
name <- gsub("TCGA-BLCA.", "", name)
colnames(countdata) <- name 
```

```{r}
name <- colnames(countdata)
name <- substr(name, 1,12)
colnames(countdata) <- name 
```

```{r}
head(colnames(countdata))
```

## Calculating TPM values for all the cohorts

```{r}
library(GenomicFeatures)
```


```{r}
name_proj <- names(x)
y <- list()
txdb <- makeTxDbFromGFF("gencode.v38.annotation.gtf")
all.genes <- genes(txdb)
```


```{r}
for (proj in name_proj) {
  TCGA_BLCA <- do.call(rbind.data.frame, x[proj])
  
  name <- colnames(TCGA_BLCA)
  name <- gsub(paste0(proj, "."), "", name)
  colnames(TCGA_BLCA) <- name
  
  name <- rownames(TCGA_BLCA)
  name <- gsub(paste0(proj, "."), "", name)
  rownames(TCGA_BLCA) <- name
  
  name <- colnames(TCGA_BLCA)
  name <- substr(name, 1,12)
  colnames(TCGA_BLCA) <- name
  
  my.genes <- rownames(TCGA_BLCA)
  exons.list.per.gene <- exonsBy(txdb,by="gene")
  exonic.gene.sizes <- as.data.frame(sum(width(reduce(exons.list.per.gene))))
  exonic.gene.sizes <- exonic.gene.sizes[my.genes,]
  names(exonic.gene.sizes) <- my.genes
  
    tpm3 <- function(counts,len) {
 
    x <- counts/len
    x <- na.omit(x)
  
    return(t(t(x)*1e6/colSums(x)))
 
    }
    
    TCGA_TPM <- as.data.frame(tpm3(as.matrix(TCGA_BLCA), exonic.gene.sizes))
    
  y[[paste0(proj)]] <- TCGA_TPM
}
```

```{r}
list.save(y, 'TCGA_onlymalemore50_TPM_ENSMBL.rdata')
```

```{r}
name_proj <- names(y)
s <- list()
for (proj in name_proj) {
  TCGA_BLCA <- do.call(rbind.data.frame, y[proj])
  
  name <- colnames(TCGA_BLCA)
  name <- gsub(paste0(proj, "."), "", name)
  colnames(TCGA_BLCA) <- name
  
  name <- rownames(TCGA_BLCA)
  name <- gsub(paste0(proj, "."), "", name)
  rownames(TCGA_BLCA) <- name
  
  name <- colnames(TCGA_BLCA)
  name <- substr(name, 1,12)
  colnames(TCGA_BLCA) <- name 
  
  symbol_count <- TCGA_BLCA

  symbol_count$symbol <- mapIds(org.Hs.eg.db,
                     keys=rownames(symbol_count),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")

  symbol_count <- aggregate(. ~ symbol, data = symbol_count, max)
  rownames(symbol_count) <- symbol_count$symbol
  
  symbol_count <- symbol_count[,-1]
  
  s[[paste0(proj)]] <- symbol_count
}
```
```{r}
list.save(s, 'TCGA_onlymalemore50_TPM_SYMBOL.rdata')
```

```{r}
counts_Y_sort <- counts_Y_new[,sort(colnames(counts_Y_new))]
```


```{r}
name_proj <- names(x)
y <- list()
for (proj in name_proj) {
  TCGA_BLCA <- do.call(rbind.data.frame, x[proj])
  
  name <- colnames(TCGA_BLCA)
  name <- gsub(paste0(proj, "."), "", name)
  colnames(TCGA_BLCA) <- name
  
  name <- rownames(TCGA_BLCA)
  name <- gsub(paste0(proj, "."), "", name)
  rownames(TCGA_BLCA) <- name
  
  name <- colnames(TCGA_BLCA)
  name <- substr(name, 1,12)
  colnames(TCGA_BLCA) <- name 
  
  symbol_count <- TCGA_BLCA

  symbol_count$symbol <- mapIds(org.Hs.eg.db,
                     keys=rownames(symbol_count),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")

  symbol_count <- aggregate(. ~ symbol, data = symbol_count, max)
  rownames(symbol_count) <- symbol_count$symbol
  
  symbol_count <- symbol_count[,-1]
  
  y[[paste0(proj)]] <- symbol_count
}
```
```{r}
list.save(y, 'TCGA_onlymalemore50_symbol.rdata')
```

```{r}
library(org.Hs.eg.db)
```

```{r}
symbol_count <- countdata
```

```{r}
symbol_count$symbol <- mapIds(org.Hs.eg.db,
                     keys=rownames(symbol_count),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```
```{r}
symbol_count <- aggregate(. ~ symbol, data = symbol_count, max)
row.names(symbol_count) <- symbol_count$symbol
```

```{r}
symbol_count <- symbol_count[,1]
```

```{r}
symbol_count <- symbol_count[,-1]
```

```{r}
library
```


```{r}
write.csv(symbol_count, "all_male_counts_symbol.cvs")
```

```{r}
for (proj in name_proj) {
  name <- colnames(symbol_count)
  tag <- paste0(proj, ".")
  name <- gsub(tag, "", name)
  colnames(symbol_count) <- name 
}
```

```{r}
name <- colnames(symbol_count)
name <- substr(name,1,12)
colnames(symbol_count) <- name 
```

```{r}
i <- which(colnames(symbol_count) %in% clinicaldata$case_submitter_id)
```

```{r}
i <- which(clinicaldata$case_submitter_id %in% colnames(symbol_count))
```

```{r}
clinicaldatacommon <- clinicaldata[i,]
```

```{r}
i <- which(colnames(symbol_count) %in% rownames(clinicaldatacommon))
```

```{r}
symbolcommon <- symbol_count[,rownames(clinicaldatacommon)]
```

```{r}
i <- which(colnames(countdata) %in% clinicaldata$case_submitter_id)
```

```{r}
name <- colnames(countdata)
name <- substr(name,1,12)
colnames(countdata) <- name 
```



```{r}
i <- which(male_ID$Patient_ID %in% rownames(LoY))
i
```

```{r}
LoY <- c(1:length(male_ID$Patient_ID))
LoY[i] <- c("Yes")
LoY[-i] <- c("No")
```

```{r}
male_ID$LoY <- LoY
```
```


```{r}
patients <- split(TCGA_BLCA, f = TCGA_BLCA$Sample)
  name <- names(patients)
  i <- which(is.duplicated(name))
  patients <- patients[-i]
```




Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
HNSC_SegmentFiles$Size <- HNSC_SegmentFiles$End - HNSC_SegmentFiles$Start
```

Look for patients that have more than one sample and cleaning it up:

```{r}
patients <- split(HNSC_SegmentFiles, f = HNSC_SegmentFiles$Sample)
```

Giving patients the same ID as other datasets:

```{r}
name <- names(patients)
```

```{r}
name <- substring(name, 1, 12)
```

```{r}
name <- gsub("-", ".", name)
```

```{r}
duplicated(name) 
```

Taking out duplicated data:

```{r}
HNSC_SegmentFiles <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample != "TCGA-CV-6955-01A-11D-2062-01",]
```

```{r}
HNSC_SegmentFiles <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample != "TCGA-CV-6956-01A-21D-2062-01",]
```

Renaming the patients with an ID matching the metadata:

```{r}
HNSC_SegmentFiles$Sample <- substring(HNSC_SegmentFiles$Sample, 1, 12)
HNSC_SegmentFiles$Sample <- gsub("-", ".", HNSC_SegmentFiles$Sample)
```

### In male patients

Selecting the male patients:

```{r}
SegmentFiles_Male <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample %in% male_ID$Patient_ID,]
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
SegmentFiles_Male$Unlog_Segment <- 2^(1+SegmentFiles_Male$Segment_Mean)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
SegmentFiles_Male$Weighted_value <- SegmentFiles_Male$Unlog_Segment*SegmentFiles_Male$Size
```

Separating the patients in a list in order for the loop to work:

```{r}
patients_male <- split(SegmentFiles_Male, f = SegmentFiles_Male$Sample )
```

Loop that calculates the average Copy Number Index per Chromosome inside each patient:

```{r}
ID <- names(patients_male)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_male))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_male[[h]]$Weighted_value[patients_male[[h]]$Chromosome %in% i])/sum(patients_male[[h]]$Size[patients_male[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each patient:

```{r}
chrom_male <- do.call(rbind.data.frame, x)
colnames(chrom_male) <- a
rownames(chrom_male) <- ID
```

Plotting the CNI for all the Male patients:

```{r}
den <- density(chrom_male$Y)
plot(den, xlim = c(0,2))
```
Obtaining the data in common patients for Percentage Tumor Nuclei and Chromosome Number Index:

```{r}
data3 <- data2[data2$ID %in% rownames(chrom_male),]
```

```{r}
chrom_male1 <- chrom_male[data3$ID,]
```

Weighting Chromosome Number Index by Percentage tumor nuclei:

```{r}
chrom_male2 <- (chrom_male1 + (data3$x/100) - 1)/(data3$x/100)
```

Saving the data:

```{r}
write.table(chrom_male2, "/Users/cristinacondel/Desktop/Gender_COX2/Male_CopyNumberIndex.txt", sep = "\t", row.names = F, quote = F)
```

Plotting weighted Y chromosome for male patients:

```{r}
den <- density(chrom_male2$`Y`)
plot(den, xlim = c(0,2), xlab = "ChrY Copy Number Index")
```

From this graph we can extract that we have successfully replicated what was done in the analysis by Hollows et al. 

Calculating the maximum points of the graph:

```{r}
which.max(den$y)
den$x[110]
```

```{r}
which.max(den$y[den$x < 0.75])
den$x[72]
```

```{r}
den <- density(chrom_male2$`X`)
plot(den)
```

```{r}
LoY <- chrom_male2[chrom_male2$Y < 0.4803115,]
```

```{r}
i <- which(male_ID$Patient_ID %in% rownames(LoY))
i
```

```{r}
LoY <- c(1:length(male_ID$Patient_ID))
LoY[i] <- c("Yes")
LoY[-i] <- c("No")
```

```{r}
male_ID$LoY <- LoY
```


### Doing the same process for the female patients:

```{r}
Female_ID <- TCGA_Masterfile[TCGA_Masterfile$Gender == "Female",]
```

```{r}
SegmentFiles_Female <- HNSC_SegmentFiles[HNSC_SegmentFiles$Sample %in% Female_ID$Patient_ID,]
```

```{r}
SegmentFiles_Female$Unlog_Segment <- 2^(1+SegmentFiles_Female$Segment_Mean)
```

```{r}
SegmentFiles_Female$Weighted_value <- SegmentFiles_Female$Unlog_Segment*SegmentFiles_Female$Size
```

```{r}
patients_Female <- split(SegmentFiles_Female, f = SegmentFiles_Female$Sample )
```

```{r}
ID <- names(patients_Female)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(patients_Female))) {
  vec <- c()
  for (i in a) {
    mean <- sum(patients_Female[[h]]$Weighted_value[patients_Female[[h]]$Chromosome %in% i])/sum(patients_Female[[h]]$Size[patients_Female[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

```{r}
chrom_Female <- do.call(rbind.data.frame, x)
```

```{r}
colnames(chrom_Female) <- a
rownames(chrom_Female) <- ID
```

```{r}
write.table(chrom_Female, "/Users/cristinacondel/Desktop/Gender_COX2/Female_CopyNumberIndex.txt", sep = "\t", row.names = T, quote = T)
```

```{r}
den <- density(chrom_Female$Y)
plot(den)
```

```{r}
den <- density(chrom_Female$X)
plot(den)
```



# Y chromosome signature expression

We now check for the expression of our Y chromosome signature overlapping with the computed status of EDY and LoY

First we stablish the signature:

```{r}
Y_sig <- c("DDX3Y",	"EIF1AY",	"KDM5D",	"NLGN4Y",	"RPS4Y1",	"SRY",	"TBL1Y",	"TMSB4Y","USP9Y",	"UTY",	"ZFY")
```

```{r}

```


Select the genes for the count table and prepare the count table to be read by the pheatmap() function:

```{r}
i <- which(rownames(Counts_symbol) %in%  Y_sig)
counts_Y <- Counts_symbol[i,]
```

```{r}
names <- rownames(counts_Y)
counts_Y <- as.data.frame(apply( counts_Y, 2, as.numeric))
rownames(counts_Y) <- names
```


```{r}
 library(pheatmap)
  choose_matrix=counts_Y
  #choose_matrix[1:4,1:4]
  choose_matrix=t(scale(t(log2(choose_matrix+1))))  
  ## http://www.bio-info-trainee.com/1980.html
  annotation_col = data.frame(LoY = EDYLoY$LoY, EDY = EDYLoY$EDY)
  rownames(annotation_col)=colnames(counts_Y)
  pheatmap(choose_matrix,show_colnames = F,annotation_col = annotation_col, show_rownames = T
           #,filename = 'CellLine_LOY_EDY_Signature_HN_SCC.png'
           )
  
```

```{r}
download.file("https://tcga-xena-hub.s3.us-east-1.amazonaws.com/download/survival%2FBLCA_survival.txt", "/Users/cristinacondel/Desktop/Gender_Analysis/Male/TCGA_allmales/survival/TCGA-BCLA-survival.txt")
```

