---
title: "EDY Analysis TCGA-HNSC"
subtitle: "Obtatining EDY status for Male Patients"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
 prettydoc::html_pretty:
    toc: true
    theme: hpstr
    highlight: github
    number_sections: true
 pdf_document:
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
bibliography: citations.bibtex
nocite: '@*'
---

# Obtaining the Data 

## Raw Counts with Symbols ID

Downloaded from the GDC using TCGABiolinks on the 2nd of December 2022:

```{r}
Sim_counts <- read.table("HNSC_COUNTS.txt")
```

Preparing the Patients ID so it matches other datasets:

```{r}
colnames(Sim_counts) <- substr(Sim_counts[1,], 1, 12)
Sim_counts <- Sim_counts[-1,]
```

```{r}
Sim_counts <- Sim_counts[,order(colnames(Sim_counts))] 
```

```{r, warning=FALSE}
rownames(Sim_counts) <- Sim_counts$Tags
Sim_counts <- Sim_counts[order(rownames(Sim_counts)),]
Sim_counts <- Sim_counts[,-1]
```

```{r}
colnames(Sim_counts) <- gsub("-", ".", colnames(Sim_counts))
```

## Patient Metadata

This data was obtained on the 7th of October downloaded directly from the GDC portal: 

```{r, warning=FALSE}
library(readxl)
TCGA_Masterfile <- read_excel("TCGA_MASTERFILE_clinical_data.xls")
TCGA_Masterfile$Patient_ID <- gsub("-", ".", TCGA_Masterfile$Patient_ID)
row.names(TCGA_Masterfile) <- TCGA_Masterfile$Patient_ID
```

## LoY Status

```{r}
male<- read.table("TCGA_Metadata_Male_Ychr.txt", sep = "\t")
colnames(male) <- male[1,]
male <- male[-1,]
```

```{r}
LoY <- as.data.frame(male$LoY)
rownames(LoY) <- male$Patient_ID
```

Introducing LoY Status into the Metadata dataset:

```{r}
TCGA_Masterfile[87] <- c(NA)
```

```{r}
colnames(TCGA_Masterfile)[87] <- "LoY"
```

```{r}
i <- which(TCGA_Masterfile$Patient_ID %in% rownames(LoY))
```

```{r}
TCGA_Masterfile$LoY[i] <- LoY$`male$LoY` 
```

# EDY status calculation

## Creating an ExpressionSet object from the Data

```{r, message=FALSE}
library(Biobase)
```

Clinical data is stored as an AnnotatedDataFrame object, this would be the "phenotypic data" in the ExpressionSet object.

```{r, warning=FALSE}
row.names(TCGA_Masterfile) <- TCGA_Masterfile$Patient_ID
```

```{r}
pData <- AnnotatedDataFrame(TCGA_Masterfile) 
```

Transforming the data frame of counts into a matrix so it's readable for the ExpressionSet() function:

```{r}
exprs <- as.matrix(Sim_counts)
```

Checking all the samples are the same in phenotypic and count data:

```{r}
all(rownames(pData) == colnames(exprs))
```

In the featureData part of the ExpressionSet object we store the **Gene Symbols** which correspond to the rownames of the counts dataframe:

```{r}
featureData <- as.data.frame(row.names(Sim_counts))
rownames(featureData) <- row.names(Sim_counts)
colnames(featureData) <- c("Symbol")
```

With all this information we create the Expression Set object:

```{r}
exampleSet <- ExpressionSet(assayData=exprs, phenoData=pData, featureData = AnnotatedDataFrame(featureData))
```

## Calculation of EDY

```{r}
#Filter males in the expression set
ii <- which(varLabels(exampleSet)=="Gender")
        exampleSet <- exampleSet[,pData(exampleSet)[,ii]=="Male"]
        exampleSet <- exampleSet[,!is.na(pData(exampleSet)[,ii])]
```

```{r}
#Add column with the hgnc symbol information to fData
fData(exampleSet)$id.feature <- featureNames(exampleSet)
```
        
```{r}
annot.expr <- fData(exampleSet)
```

```{r}
gene.expr <- exprs(exampleSet)[rownames(exprs(exampleSet))%in%annot.expr$id.feature,]
      #Replace gene ID for hgnc symbol
      rownames(gene.expr) <- annot.expr[, "Symbol"]
      #Select those genes that belong to chrY
      exprY <- gene.expr[annot.expr[, "Symbol"]%in%EDY::chrY.genes$hgnc_symbol,]
      exprY <- exprY[complete.cases(exprY),]
      #Select those genes that belong to the rest of the genome
      exprRef <- gene.expr[annot.expr[, "Symbol"]%in%EDY::autosomal.genes$hgnc_symbol,]
      exprRef <- exprRef[complete.cases(exprRef),]
```


```{r}
#Convert expression sets into numeric since they came out as character
exprY <- as.data.frame(apply( exprY, 2, as.numeric))
exprRef <- as.data.frame(apply(exprRef, 2, as.numeric))
```

```{r}
#Apply EDY formulae: 
Ry <- sweep(log2(exprY+1), 2, FUN="-", 
                apply(log2(exprRef+1), 2, mean))
EDYcontinuous <- apply(Ry, 2, mean)
```

```{r}
# Establish the control group, in this case males known for having no LoY
controls <- EDYcontinuous[pData(exampleSet)[,"LoY"]=="No"]
```

```{r}
# Establish from the cutoff value cases with EDY
thresh <- median(controls, na.rm=TRUE) - 1.2*IQR(controls, na.rm=TRUE)
  EDY <- cut(EDYcontinuous, c(-Inf, thresh, Inf), 
             labels = c("Yes", "No"))
  EDY <- relevel(EDY, 2)
```

```{r}
# Build the output
pData(exampleSet)$EDY <- EDY
    names(EDY) <- names(EDYcontinuous)
    ans <- list(EDY=EDY, EDYcontinuous=EDYcontinuous, 
                threshold=thresh)
```

```{r}
table(ans$EDY)
```

Creating a dataset from the list obtained in the result:

```{r}
ans1 <- do.call(rbind.data.frame, ans)
```

```{r}
colnames(ans1) <- male$Patient_ID
rownames(ans1) <- c("EDY", "EDYcontinuous", "Threshold")
```

# Saving the data

```{r}
write.table(ans1, "/Users/cristinacondel/Desktop/Gender_COX2/Male_HNSC_EDYinformation.txt", sep = "\t", row.names = F, quote = F)
```

```{r}
TCGA_Masterfile[88] <- c(NA)
```

```{r}
colnames(TCGA_Masterfile)[88] <- "EDY"
```

```{r}
i <- which(TCGA_Masterfile$Patient_ID %in% names(ans$EDY))
```

```{r}
TCGA_Masterfile$EDY[i] <- factor(ans$EDY, levels = c("No","Yes"), labels = c(1, 2))
```

```{r}
table(TCGA_Masterfile[TCGA_Masterfile$Gender == "Male",]$LoY, TCGA_Masterfile[TCGA_Masterfile$Gender == "Male",]$EDY)
```

# Bibliography

