---
title: "CNV Analysis CCLine for SCC"
subtitle: "Obtatining Copy Number Index"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
 prettydoc::html_pretty:
    toc: true
    theme: hpstr
    highlight: github
    number_sections: true
 pdf_document:
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---

# Obtain the Data from cBioportal

## CNV data

This data was downloaded on the 22nd of February 2022 from the Xenabrowser using R tool download.file from the following link: 
https://data.broadinstitute.org/ccle_legacy_data/dna_copy_number/CCLE_copynumber_2013-12-03.seg.txt.

```{r}
Segment_cell <- read.table("cBiosegmentdata.txt", sep = "\t" )
```

## Clinical data

This data was downloaded on the 22nd of February 2022 from the Xenabrowser using R tool download.file from the following link: 
https://data.broadinstitute.org/ccle_legacy_data/cell_line_annotations/CCLE_sample_info_file_2012-10-18.txt.

```{r}
Clinical_cell <- read.table("cBioclinicaldata.txt", sep = "\t" )
```

From the same portal we generate a file which has the names for the Upper Aerodigestive Track samples:

```{r}
hn_samples <- read.table("hnn_sample.txt", sep = "\t" )
```

## Selecting samples of interest

We select those samples that are of interest:

```{r}
# Which are Upper Aerodigestive Track samples
i <- which(Clinical_cell$`CCLE name` %in% hn_samples$V1)
hn_clinical <- Clinical_cell[i,] 
```

```{r}
# Select the male samples which come from squamous cell carcinoma
hn_cell_female <- hn_clinical[hn_clinical$Gender == "F",]
hn_cell_female <- hn_cell_female[hn_cell_female$`Hist Subtype1` == "squamous_cell_carcinoma",]
```


```{r}
# Select the segmented data for this samples
i <- which(Segment_cell$CCLE_name %in% hn_cell_female$`CCLE name`)
hn_segmented <- Segment_cell[i,] 
```

```{r}
colnames(hn_segmented) <- colnames(Segment_cell)
```

# Processing of the data:

## Obtaining CNI

Create Sizes for the segments: Chromosome End - Chromosome Star:

```{r}
hn_segmented[3:4] <- lapply(hn_segmented[3:4], as.numeric)
```

```{r}
hn_segmented$Size <- hn_segmented$End - hn_segmented$Start
```

Unlogging the segment_mean variable to obtain Copy Number:

```{r}
hn_segmented[6] <- lapply(hn_segmented[6], as.numeric)
```

```{r}
hn_segmented$Unlog_Segment <- 2^(1+hn_segmented$Segment_Mean)
```

Weighting the segment_mean by the sizes of the segments:

```{r}
hn_segmented$Weighted_value <- hn_segmented$Unlog_Segment*hn_segmented$Size
```

Separating the samples in a list in order for the loop to work:

```{r}
cells <- split(hn_segmented, f = hn_segmented$CCLE_name)
```

Loop that calculates the average Copy Number Index per Chromosome inside each sample:

```{r}
ID <- names(cells)
a <- c(1:22, "X", "Y")
x <- list()
for (h in (1:length(cells))) {
  vec <- c()
  for (i in a) {
    mean <- sum(cells[[h]]$Weighted_value[cells[[h]]$Chromosome %in% i])/sum(cells[[h]]$Size[cells[[h]]$Chromosome %in% i])
    vec <- c(vec,mean)
  }
  names(vec) <- c(1:22, "X", "Y")
  x[[paste0(ID[h])]] <- vec
}

```

Unlisting the result to obtain a data frame with the CNI for each Chromosome for each sample:

```{r}
chrom_cells <- do.call(rbind.data.frame, x)
colnames(chrom_cells) <- a
rownames(chrom_cells) <- ID
```

Plotting the Y chromosome index for all samples:

```{r}
den <- density(chrom_cells$Y)
plot(den, xlim = c(0,2))
```

```{r}
den <- density(chrom_cells$X)
plot(den)
```
