---
title: "EDY Analysis CCLine for SCC"
subtitle: "Stablishing EDY status"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
 prettydoc::html_pretty:
    toc: true
    theme: hpstr
    highlight: github
    number_sections: true
 pdf_document:
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---

# Obtain the Data from XenaBrowser

## Raw count data

This data was downloaded on the 24th of February 2022 from the Xenabrowser using R tool fread() from the following link: 
https://data.broadinstitute.org/ccle/CCLE_DepMap_18Q2_RNAseq_reads_20180502.gct.

```{r, message=FALSE}
library(readr)
cell_counts <- read_csv("cell_counts.cvs")
```

Preparing the raw count data so it's readable for the pipeline

```{r}
cell_counts <- aggregate(. ~ Description, data = cell_counts, max)
cell_counts1 <- cell_counts[,c(-1,-2)]
cell_counts1 <- as.data.frame(cell_counts1)
```

## Clinical data

We upload the curated metadata info extrated from the LoY pipeline:

```{r, message=FALSE, warning=FALSE}
LoY <- read_csv("LoY_cell_info.cvs")
rownames(LoY) <- LoY$CCLE_Name
```

We extract the selected samples from the count file:

```{r}
i <- which(colnames(cell_counts1) %in% rownames(LoY))
cell_counts_HN <- cell_counts1[,i]
```

```{r}
rownames(cell_counts_HN) <- cell_counts$Description
```

# EDY status calculation

## Creating an ExpressionSet object from the Data

```{r, message=FALSE}
library(Biobase)
```

There's one sample for which there's no count data, we check the name of the sample and take it out of the clinical dataset so the pipeline works.

```{r}
LoY[10,]
```

Clinical data is stored as an AnnotatedDataFrame object, this would be the "phenotypic data" in the ExpressionSet object. In this case we need to 

```{r}
pData <- AnnotatedDataFrame(LoY[-10,])
rownames(pData) <- LoY$CCLE_Name[-10]
```

Transforming the data frame of counts into a matrix so it's readable for the ExpressionSet() function:

```{r}
exprs <- as.matrix(cell_counts_HN)
```

Checking all the samples are the same in phenotypic and count data:

```{r}
all(rownames(pData) == colnames(exprs))
```

In the featureData part of the ExpressionSet object we store the **Gene Symbols** which correspond to the rownames of the counts dataframe:

```{r}
featureData <- as.data.frame(row.names(cell_counts_HN))
rownames(featureData) <- row.names(cell_counts_HN)
colnames(featureData) <- c("Symbol")
```

With all this information we create the Expression Set object:

```{r}
exampleSet <- ExpressionSet(assayData=exprs, phenoData=pData, featureData = AnnotatedDataFrame(featureData))
```

## Calculation of EDY


```{r}
#Add column with the hgnc symbol information to fData
fData(exampleSet)$id.feature <- featureNames(exampleSet)
```
        
```{r}
annot.expr <- fData(exampleSet)
```

```{r}
gene.expr <- exprs(exampleSet)[rownames(exprs(exampleSet))%in%annot.expr$id.feature,]
      #Replace gene ID for hgnc symbol
      rownames(gene.expr) <- annot.expr[, "Symbol"]
      #Select those genes that belong to chrY
      exprY <- gene.expr[annot.expr[, "Symbol"]%in%EDY::chrY.genes$hgnc_symbol,]
      exprY <- exprY[complete.cases(exprY),]
      #Select those genes that belong to the rest of the genome
      exprRef <- gene.expr[annot.expr[, "Symbol"]%in%EDY::autosomal.genes$hgnc_symbol,]
      exprRef <- exprRef[complete.cases(exprRef),]
```


```{r}
#Convert expression sets into numeric since they came out as character
exprY <- as.data.frame(apply( exprY, 2, as.numeric))
exprRef <- as.data.frame(apply(exprRef, 2, as.numeric))
```

```{r}
#Apply EDY formulae: 
Ry <- sweep(log2(exprY+1), 2, FUN="-", 
                apply(log2(exprRef+1), 2, mean))
EDYcontinuous <- apply(Ry, 2, mean)
```

```{r}
# Establish the control group, in this case males known for having no LoY
controls <- EDYcontinuous[pData(exampleSet)[,"LoY"]=="No"]
```

```{r}
# Establish from the cutoff value cases with EDY
thresh <- median(controls, na.rm=TRUE) - 1.2*IQR(controls, na.rm=TRUE)
  EDY <- cut(EDYcontinuous, c(-Inf, thresh, Inf), 
             labels = c("Yes", "No"))
  EDY <- relevel(EDY, 2)
```

```{r}
# Build the output
pData(exampleSet)$EDY <- EDY
    names(EDY) <- names(EDYcontinuous)
    ans <- list(EDY=EDY, EDYcontinuous=EDYcontinuous, 
                threshold=thresh)
```

```{r}
EDY <- ans$EDY
```

```{r}
EDYLoY <- cbind(LoY[-10,], EDY)
```

```{r}
table(EDYLoY$EDY)
```

```{r}
table(EDYLoY$LoY)
```

```{r}
table(EDYLoY$LoY, EDYLoY$EDY)
```

```{r}
write_csv(EDYLoY, "Cell_EDYLoY.cvs")
```

Creating a dataset from the list obtained in the result:

```{r}
ans1 <- do.call(rbind.data.frame, ans)
```

```{r}
colnames(ans1) <- colnames(cell_counts_HN)
rownames(ans1) <- c("EDY", "EDYcontinuous", "Threshold")
```

# Y chromosome signature expression

We now check for the expression of our Y chromosome signature overlapping with the computed status of EDY and LoY

First we stablish the signature:

```{r}
Y_sig <- c("DDX3Y",	"EIF1AY",	"KDM5D",	"NLGN4Y",	"RPS4Y1",	"SRY",	"TBL1Y",	"TMSB4Y","USP9Y",	"UTY",	"ZFY")
```

Select the genes for the count table and prepare the count table to be read by the pheatmap() function:

```{r}
i <- which(rownames(cell_counts_HN) %in%  Y_sig)
counts_Y <- cell_counts_HN[i,]
```

```{r}
USP9Y <- cell_counts_HN[c("USP9Y"),]
counts_Y <- rbind(counts_Y, USP9Y)
```

```{r}
names <- rownames(counts_Y)
counts_Y <- as.data.frame(apply( counts_Y, 2, as.numeric))
rownames(counts_Y) <- names
```


```{r}
annoCol<-list(LoY= c(No= "palegreen2", Yes = "palevioletred1"), EDY = c(No = "skyblue1", Yes ="sandybrown"))
 library(pheatmap)
  choose_matrix=counts_Y
  #choose_matrix[1:4,1:4]
  choose_matrix=t(scale(t(log2(choose_matrix+1))))  
  ## http://www.bio-info-trainee.com/1980.html
  annotation_col = data.frame(LoY = EDYLoY$LoY, EDY = EDYLoY$EDY)
  rownames(annotation_col)=colnames(counts_Y)
  pheatmap(choose_matrix,show_colnames = F,annotation_col = annotation_col, show_rownames = T, annotation_colors = annoCol
           ,filename = 'CellLine_LOY_EDY_Signature_HN_SCC.png'
           )
  
```



