---
title: "EDY Analysis CPTAC-HNSCC"
subtitle: "Stablishing EDY status"
author: "Cristina Conde López"
date: "`r Sys.Date()`"
output: 
 prettydoc::html_pretty:
    toc: true
    theme: hpstr
    highlight: github
    number_sections: true
 pdf_document:
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---

# Obtain the Data from TCGABiolinks

## Raw count data

This data was downloaded on the 2nd of March 2022 using the different tools provided by TCGABiolinks to download data from GDC Portal:

Loading of package TCGAbiolink (download raw data)

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("TCGAbiolinks")
```

```{r}
library (TCGAbiolinks)
```

In this case selecting head and neck cancer, transcriptome profiling data --> gene expression quatification, data type in counts.

```{r eval=FALSE}
query <- GDCquery(project = 'CPTAC-3',
data.category = "Transcriptome Profiling",
data.type = "Gene Expression Quantification",
workflow.type = "HTSeq - Counts", sample.type = "Primary Tumor")
GDCdownload(query)
```

GDCprepare converts the data downloaded into an R file

```{r eval=FALSE}
CPTAC_HTseqSE <- GDCprepare(query, save = TRUE, save.filename =
"CPTAC_3_HTseq.rda")
```

Extracting the counts (under assay) using SummarizedExperiment from the HNSC_HTseqSE object

```{r eval=FALSE}
Counts <- as.data.frame(SummarizedExperiment::assay(CPTAC_HTseqSE))
```

Writing the table into the working directory:

```{r eval=FALSE}
names <- substr(colnames(Counts), 1, 9)
colnames(Counts) <- names
write.csv(as.data.frame(Counts), 
          file="/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_3_counts.csv")
```

```{r}
Counts <- read.csv(file="/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_3_counts.csv")
```

```{r}
rownames(Counts) <- Counts$X
Counts <- Counts[,c(-1)]
```

```{r}
names <- substr(colnames(Counts), 1, 9)
colnames(Counts) <- names
```

```{r}

```

```{r}
library(org.Hs.eg.db)
```

```{r}
Counts$symbol <- mapIds(org.Hs.eg.db,
                     keys=rownames(Counts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
Counts_symbol <- aggregate(x = Counts[ , colnames(Counts) != "symbol"],             # Mean by group
          by = list(Counts$symbol),
          FUN = max)
row.names(Counts_symbol) <- Counts_symbol$symbol
```

```{r}
names <- substr(colnames(Counts_symbol),1,9) 
```

```{r}
colnames(Counts_symbol) <- names
```

```{r}
Counts_symbol <- Counts_symbol[,c(-1,-2)]
```

## Clinical data

We upload the curated metadata info extrated from the LoY pipeline:

```{r, message=FALSE, warning=FALSE}
LoY <- read.csv("/Users/cristinacondel/Desktop/Gender_Analysis/Male/CPTAC/CPTAC_clinical_HNSSC_male_LoY.csv")
```

```{r}
LoY$submitter_id <- gsub("-", ".", LoY$submitter_id)
```

We extract the selected samples from the count file:

```{r}
i <- which(colnames(Counts) %in% LoY$submitter_id)
Counts_male <- Counts[,i]
```

```{r}
Counts_male <- Counts_male[, sort(colnames(Counts_male))]
```


```{r}
i <- which(LoY$submitter_id %in% colnames(Counts))
LoY_overlap <- LoY[i,]
rownames(LoY_overlap) <- LoY_overlap$submitter_id
```

```{r}
LoY_overlap <- LoY_overlap[sort(rownames(LoY_overlap)),]
```



# EDY status calculation

## Creating an ExpressionSet object from the Data

```{r, message=FALSE}
library(Biobase)
```

Clinical data is stored as an AnnotatedDataFrame object, this would be the "phenotypic data" in the ExpressionSet object. In this case we need to 

```{r}
pData <- AnnotatedDataFrame(LoY_overlap)
rownames(pData) <- LoY_overlap$submitter_id
```

Transforming the data frame of counts into a matrix so it's readable for the ExpressionSet() function:

```{r}
library(org.Hs.eg.db)
```

```{r}
Counts_male$symbol <- mapIds(org.Hs.eg.db,
                     keys=rownames(Counts_male),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
Counts_symbol <- aggregate(. ~ symbol, data = Counts_male, max)
row.names(Counts_symbol) <- Counts_symbol$symbol
```

```{r}
Counts_symbol <- Counts_symbol[,c(-1)]
```

```{r}
exprs <- as.matrix(Counts_symbol)
```

Checking all the samples are the same in phenotypic and count data:


```{r}
all(rownames(pData) == colnames(exprs))
```

In the featureData part of the ExpressionSet object we store the **Gene Symbols** which correspond to the rownames of the counts dataframe:

```{r}
featureData <- as.data.frame(row.names(Counts_symbol))
rownames(featureData) <- row.names(Counts_symbol)
colnames(featureData) <- c("Symbol")
```

With all this information we create the Expression Set object:

```{r}
exampleSet <- ExpressionSet(assayData=exprs, phenoData=pData, featureData = AnnotatedDataFrame(featureData))
```

## Calculation of EDY


```{r}
#Add column with the hgnc symbol information to fData
fData(exampleSet)$id.feature <- featureNames(exampleSet)
```
        
```{r}
annot.expr <- fData(exampleSet)
```

```{r}
gene.expr <- exprs(exampleSet)[rownames(exprs(exampleSet))%in%annot.expr$id.feature,]
      #Replace gene ID for hgnc symbol
      rownames(gene.expr) <- annot.expr[, "Symbol"]
      #Select those genes that belong to chrY
      exprY <- gene.expr[annot.expr[, "Symbol"]%in%EDY::chrY.genes$hgnc_symbol,]
      exprY <- exprY[complete.cases(exprY),]
      #Select those genes that belong to the rest of the genome
      exprRef <- gene.expr[annot.expr[, "Symbol"]%in%EDY::autosomal.genes$hgnc_symbol,]
      exprRef <- exprRef[complete.cases(exprRef),]
```


```{r}
#Convert expression sets into numeric since they came out as character
exprY <- as.data.frame(apply( exprY, 2, as.numeric))
exprRef <- as.data.frame(apply(exprRef, 2, as.numeric))
```

```{r}
#Apply EDY formulae: 
Ry <- sweep(log2(exprY+1), 2, FUN="-", 
                apply(log2(exprRef+1), 2, mean))
EDYcontinuous <- apply(Ry, 2, mean)
```

```{r}
# Establish the control group, in this case males known for having no LoY
controls <- EDYcontinuous[pData(exampleSet)[,"LoY"]=="No"]
```

```{r}
controls <- EDYcontinuous
```


```{r}
# Establish from the cutoff value cases with EDY
thresh <- median(controls, na.rm=TRUE) - 1.2*IQR(controls, na.rm=TRUE)
  EDY <- cut(EDYcontinuous, c(-Inf, thresh, Inf), 
             labels = c("Yes", "No"))
  EDY <- relevel(EDY, 2)
```

```{r}
# Build the output
pData(exampleSet)$EDY <- EDY
    names(EDY) <- names(EDYcontinuous)
    ans <- list(EDY=EDY, EDYcontinuous=EDYcontinuous, 
                threshold=thresh)
```

```{r}
EDY <- ans$EDY
```

```{r}
EDYLoY <- cbind(LoY_overlap, EDY)
```

```{r}
table(EDYLoY$EDY)
```

```{r}
table(EDYLoY$LoY)
```

```{r}
table(EDYLoY$LoY, EDYLoY$EDY)
```

```{r}
#write.csv(EDYLoY, "Cell_EDYLoY.cvs")
```

Creating a dataset from the list obtained in the result:

```{r}
ans1 <- do.call(rbind.data.frame, ans)
```

```{r}
colnames(ans1) <- colnames(Counts_symbol)
rownames(ans1) <- c("EDY", "EDYcontinuous", "Threshold")
```

```{r}
ans1 <- rbind(ans1, LoY_overlap$LoY)
```


# Y chromosome signature expression

We now check for the expression of our Y chromosome signature overlapping with the computed status of EDY and LoY

First we stablish the signature:

```{r}
Y_sig <- c("DDX3Y",	"EIF1AY",	"KDM5D",	"NLGN4Y",	"RPS4Y1",	"SRY",	"TBL1Y",	"TMSB4Y","USP9Y",	"UTY",	"ZFY")
```

Select the genes for the count table and prepare the count table to be read by the pheatmap() function:

```{r}
i <- which(rownames(Counts_symbol) %in%  Y_sig)
counts_Y <- Counts_symbol[i,]
```

```{r}
names <- rownames(counts_Y)
counts_Y <- as.data.frame(apply( counts_Y, 2, as.numeric))
rownames(counts_Y) <- names
```

```{r}
names <- gsub("-", ".", LoY$submitter_id)
LoY$submitter_id <- names
```


```{r}
 library(pheatmap)
  choose_matrix=counts_Y
  #choose_matrix[1:4,1:4]
  choose_matrix=t(scale(t(log2(choose_matrix+1))))  
  ## http://www.bio-info-trainee.com/1980.html
  annotation_col = data.frame(LoY = EDYLoY$LoY, EDY = EDYLoY$EDY)
  rownames(annotation_col)=colnames(counts_Y)
  pheatmap(choose_matrix,show_colnames = F,annotation_col = annotation_col, show_rownames = T
           #,filename = 'CellLine_LOY_EDY_Signature_HN_SCC.png'
           )
  
```

Select the genes for the count table and prepare the count table to be read by the pheatmap() function:

```{r}
i <- which(rownames(Counts_symbol) %in%  EDY::chrY.genes$hgnc_symbol)
counts_Y_total <- Counts_symbol[i,]
```

```{r}
names <- rownames(counts_Y_total)
counts_Y_total <- as.data.frame(apply( counts_Y_total, 2, as.numeric))
rownames(counts_Y_total) <- names
```

```{r}
nrow(counts_Y_total)
keep <- rowSums(counts_Y_total)  > 0
counts_Y_total <- as.data.frame(counts_Y_total[keep,])
nrow(counts_Y_total)
```


```{r}
 library(pheatmap)
  choose_matrix=counts_Y_total
  #choose_matrix[1:4,1:4]
  choose_matrix=t(scale(t(log2(choose_matrix+1))))  
  ## http://www.bio-info-trainee.com/1980.html
  annotation_col = data.frame(LoY = EDYLoY$LoY, EDY = EDYLoY$EDY)
  rownames(annotation_col)=colnames(counts_Y_total)
  pheatmap(choose_matrix,show_colnames = F,annotation_col = annotation_col, show_rownames = T
           #,filename = 'CellLine_LOY_EDY_Signature_HN_SCC.png'
           )
  
```

```{r}
 library(pheatmap)
  choose_matrix=counts_Y_total
  #choose_matrix[1:4,1:4]
  choose_matrix=t(scale(t(log2(choose_matrix+1))))  
  ## http://www.bio-info-trainee.com/1980.html
  clustering <- hclust(dist(t(choose_matrix), method = "euclidean"), method = "ward.D2")
  ## http://www.bio-info-trainee.com/1980.html
  annotation_col = data.frame(LoY = EDYLoY$LoY, EDY = EDYLoY$EDY)
  rownames(annotation_col)=colnames(counts_Y_total)
  pheatmap(choose_matrix,show_colnames = F,annotation_col = annotation_col, show_rownames = T, cluster_cols = clustering, cutree_cols = 3 
           #,filename = 'CellLine_LOY_EDY_Signature_HN_SCC.png'
           )
  
```


```{r}
 library(pheatmap)
  choose_matrix=counts_Y
  #choose_matrix[1:4,1:4]
  choose_matrix=t(scale(t(log2(choose_matrix+1))))
  clustering <- hclust(dist(t(choose_matrix), method = "euclidean"), method = "ward.D2")
  ## http://www.bio-info-trainee.com/1980.html
  annotation_col = data.frame(LoY = EDYLoY$LoY, EDY = EDYLoY$EDY)
  rownames(annotation_col)=colnames(counts_Y)
  pheatmap(choose_matrix,show_colnames = F,annotation_col = annotation_col, show_rownames = T, , cluster_cols = clustering, cutree_cols = 2 
           #,filename = 'CellLine_LOY_EDY_Signature_HN_SCC.png'
           )
  
```

We introduce the data in the Patient Metadata table:
```{r}
t1 <- cutree(clustering, 2)
```

```{r}
EDYLoY$Y_expres <- factor(t1, levels = c(1,2), labels = c("Low", "High"))
```

We check by size we have named the groups correctly

```{r}
table(EDYLoY$Y_expres)
```

```{r}
View(data.frame(EDYLoY$submitter_id, EDYLoY$Y_expres))
```

We plot the survival curve based on the groups:
TCGA_Masterfile_male <- transform(TCGA_Masterfile_male, OS_5y_months = as.numeric(OS_5y_months))
library(survival)
library(survminer)
surv_object0 <- Surv(time = TCGA_Masterfile_male$OS_5y_months, event = as.double(TCGA_Masterfile_male$OS_5y_event))
fit00 <- survfit(surv_object0 ~ Y_expres, data = TCGA_Masterfile_male)
ggsurvplot(fit00, data = TCGA_Masterfile_male, pval = TRUE, title = "OS 5 year event")
```{r}

```
